---
title: "RP #2"
author: "Mason Mazurek"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rp2 <- read.csv("sds358proj.csv")
mydata <- rp2[c(1:18)]
library(tidyverse)
library(matrixStats)
```

## Data Snippet

```{r show, layout="l-body-outset"}
library(knitr)
kable(head(mydata))
```

## Exploratory Data Analysis

# Correlation Table
```{r cor}
cor(mydata[3:18])
```

# Mean and SD's
```{r mean}
colMeans(mydata[3:18])
```
```{r sd}
apply(mydata[3:18], 2, sd)
```
```{r range}
apply(mydata[3:18], 2, range)
```
# Plots:

```{r w, echo=FALSE}
ggplot(rp2, aes(x=W)) +geom_histogram(bins = 10) + labs(title ="Distribution of Wins")
```

Unimodal, mean 80, range 40-110

```{r ERA, echo=FALSE}
ggplot(rp2, aes(x=ERA)) +geom_histogram(bins = 10) + labs(title ="Distribution of Earned Runs Allowed")
```
Unimodal, Mean 4, Range 2.5-5.5

```{r SV, echo=FALSE}
ggplot(rp2, aes(x=SV)) +geom_histogram(bins = 10) + labs(title ="Distribution of Saves credited to Pitcher")
```
Unimodal, Mean 40, Range 20-70

```{r IP, echo=FALSE}
ggplot(rp2, aes(x=IP)) +geom_histogram(bins = 10) + labs(title ="Distribution of Innings Pitched By Pitcher")
```
Unimodal, Slight Left Skew, Range 1410-1495, M ean 1445

```{r H, echo=FALSE}
ggplot(rp2, aes(x=H)) +geom_histogram(bins = 10) + labs(title ="Distribution of Hits allowed")
```
Unimodal, Mean 1400, Range 1100-1700

```{r R, echo=FALSE}
ggplot(rp2, aes(x=R)) +geom_histogram(bins = 10) + labs(title ="Distribution of Runs allowed")
```
Unimodal Slight Left Skew, Range 500-900, Mean 700

```{r ER, echo=FALSE}
ggplot(rp2, aes(x=ER)) +geom_histogram(bins = 10) + labs(title ="Distribution of Earned Runs allowed")
```
Unimodal, Mean 640, Range 450-900

```{r HR, echo=FALSE}
ggplot(rp2, aes(x=HR)) +geom_histogram(bins = 10) + labs(title ="Distribution of Home Runs allowed")
```
Unimodal, Slight Left Skew, Mean 175, Range 100-275

```{r BB, echo=FALSE}
ggplot(rp2, aes(x=BB)) +geom_histogram(bins = 10) + labs(title ="Distribution of Base Walks")
```
Unimodal, Mean 480, Range 300-700

```{r SO, echo=FALSE}
ggplot(rp2, aes(x=SO)) +geom_histogram(bins = 10) + labs(title ="Distribution of Strikeouts")
```
Unimodal, Mean 1275, Range 1000-1700

```{r AVG, echo=FALSE}
ggplot(rp2, aes(x=AVG)) +geom_histogram(bins = 10) + labs(title ="Distribution of Average Hits Per Game")
```
Unimodal, Mean .250, range .2-.3

```{r AB, echo=FALSE}
ggplot(rp2, aes(x=AB)) +geom_histogram(bins = 10) + labs(title ="Distribution of At Bats")
```
Unimodal, Very slight Right Skew, Mean 5500, Range 5300-5700

```{r RBI, echo=FALSE}
ggplot(rp2, aes(x=RBI)) +geom_histogram(bins = 10) + labs(title ="Distribution of Runs from Batting")
```
Unimodal, Mean 675, Range 500-900

```{r CS, echo=FALSE}
ggplot(rp2, aes(x=CS)) +geom_histogram(bins = 10) + labs(title ="Distribution of Put Outs")
```
Unimodal, Slight Left Skew, Range 10-70, Mean 35

```{r OBP, echo=FALSE}
ggplot(rp2, aes(x=OBP)) +geom_histogram(bins = 10) + labs(title ="Distribution of On-Base Percentages")
```
Unimodal, Slight Left Skew, Mean = .315, Range .29-.35

```{r SLG, echo=FALSE}
ggplot(rp2, aes(x=SLG)) +geom_histogram(bins = 10) + labs(title ="Distribution of Slugging Percentage")
```
Unimodal, Mean .4, Range .34-.46